A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-19 tex/codatadp.tex
moved around

\paragraph{Phase 2: Applying Acyclicity and Bisimilarity.}
%For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
The rules of the second phase contain premises that involve a mapping $\Val$
that associates a potentially cyclic $\mu$-term as representative with each equivalence class.
Recall that $\mu$-bindings can be used for representing possibly cyclic terms and values.
%(Section~\ref{sec:introduction}).
% for example, $\mu x. \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
Formally, $\mu$-terms are defined recursively as being either a variable $x$
or a node of the form
$\mu x.\; \const{C}( \vec t\, )$ for some constructor $\const{C} \in \Ctr$ and
$\mu$-terms $\vec t$ with the expected types.
The variable $x$ need not occur under the binder's body, in which case the
binder can be omitted.

@@

Intuitively, a $\mu$-term $t$ of type $\tau$ describes a class of values of $\tau$,
%either in the case when $\tau$ is a codatatype or datatype type,
where a cyclic $t$ describes an infeasible class of values when
%$\tau \not\in \Codata$.
$\tau \in \Data$.
At the beginning of this phase, 
we compute a mapping $\Val$ from equivalence classes to $\mu$-terms 
that describe the class of values that terms in that equivalence class can take in the models of $\Ec$. 
For each equivalence class $\ec{x}$, we associate a fresh variable $v_{\ec{x}}$ of the same type as $x$ not occurring in $\tEc$,
and initially say that $\Val$ contains $\ec{x} \mapsto v_{\ec{x}}$,
in other words, there are initially no constraints on the values for any equivalence class $\ec{x}$.
Then, $\Val$ is modified by applying the following update exhaustively:
\[
\inferrule{
  v_{\ec{x}} \in \FV( \Val )
  \quad
  \const{C}( t_1, \ldots, t_n ) \in \ec{x}
  \quad
  \const{C} \in \Ctr
}{
  \Val := \Val \{ v_{\ec{x}} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{t_1}}, \ldots, v_{\ec{t_n}} ) \}
}
\]
We write $\FV( \Val )$ to denote the \relax{free variables} occurring in $\Val$, which
exclude variables that occur beneath $\mu$-binders.
It is straightforward to see that the height of terms produced as a result of this update
is bounded by the number of equivalence classes of $\Ec$,
and thus our construction of $\Val$ is terminating.

>>>>

\paragraph{Phase 2: Applying Acyclicity and Bisimilarity.}
%For presentation of the rules in this phase, we rely on a representation of terms in the $\mu$-notation.
The rules of the second phase contain premises that involve a mapping $\Val$
that associates a $\mu$-term as representative with each equivalence class.
%Recall that $\mu$-bindings can be used for representing possibly cyclic terms and values.
%(Section~\ref{sec:introduction}).
% for example, $\mu x. \const{C}( \const{0}, x )$ represents the cyclic value $\const{C}( \const{0}, \const{C}( \const{0}, \ldots ))$.
Intuitively, a $\mu$-term $t$ of type $\tau$ describes a class of values of $\tau$,
%either in the case when $\tau$ is a codatatype or datatype type,
where a cyclic $t$ describes an infeasible class of values when
%$\tau \not\in \Codata$.
$\tau \in \Data$.
Formally, $\mu$-terms are defined recursively as being either a variable $x$
or a node of the form
$\mu x.\; \const{C}( \vec t\, )$ for some constructor $\const{C} \in \Ctr$ and
$\mu$-terms $\vec t$ with the expected types.
The variable $x$ need not occur under the binder's body, in which case the
binder can be omitted.

@@

At the beginning of this phase, 
we compute a mapping $\Val$ from equivalence classes to $\mu$-terms 
that describe the class of values that terms in that equivalence class can take in the models of $\Ec$. 
For each equivalence class $\ec{x}$, we associate a fresh variable $v_{\ec{x}}$ of the same type as $x$ not occurring in $\tEc$,
and initially say that $\Val$ contains $\ec{x} \mapsto v_{\ec{x}}$,
in other words, there are initially no constraints on the values for any equivalence class $\ec{x}$.
Then, $\Val$ is modified by applying the following update exhaustively:
\[
\inferrule{
  v_{\ec{x}} \in \FV( \Val )
  \quad
  \const{C}( t_1, \ldots, t_n ) \in \ec{x}
  \quad
  \const{C} \in \Ctr
}{
  \Val := \Val \{ v_{\ec{x}} \mapsto \mu v_{\ec{x}}. \const{C}( v_{\ec{t_1}}, \ldots, v_{\ec{t_n}} ) \}
}
\]
We write $\FV( \Val )$ to denote the \relax{free variables} occurring in $\Val$, which
exclude variables that occur beneath $\mu$-binders.
It is straightforward to see that the height of terms produced as a result of this update
is bounded by the number of equivalence classes of $\Ec$,
and thus our construction of $\Val$ is terminating.
